---
title: "Practical Machine Learning Assignment 4"
author: "Igor Gamayun"
date: "7/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Background  

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data Source  

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


## Data Processing 
1.Libraries needed for the data processing
```{r,echo=TRUE, results='hide', message=FALSE}
#loading libraries
library('knitr')
library('caret')
```

2.Creating variables for training and test data
```{r,echo=TRUE, results='hide', message=FALSE}
# getting data into variables
TrainData<-read.csv("pml-training.csv",sep = ',')
TestData<-read.csv("pml-testing.csv",sep = ',')

```
3.Cleaning training data 
```{r,echo=TRUE, results='hide', message=FALSE}
data<-data.frame(TrainData)
# choose columns without na
data1<-data[,colSums(is.na(data))==0]
# choose columns without empty cells
data1<-data1[,colSums(data1=='')==0]

# make the outcome character variable "classe" as factor variable
data1$classe=as.factor(data1$classe)
```
For this particular data character variable user_name can be included as factor variable. However, to train more general model, we will exclude user_name along with other non-numerical variables (time...etc) from the analysis. Therefore all the model will be trained only on the data from movement sensors, which we will set as numeric variables.
```{r,echo=TRUE, results='hide', message=FALSE}
#extra variable we can compare two conditions: with the name as a variable and without
data1$user_name=as.factor(data1$user_name)
dataTrain1=data1[,-c(1:7)];

dataTrain1[,-53]=sapply(dataTrain1[,-53],as.numeric)
```
## creating data partition
```{r,echo=TRUE, results='hide', message=FALSE}
#create data partition
set.seed(12345)
index<-createDataPartition(dataTrain1$classe,p=0.75, list=FALSE)
dataTrain=dataTrain1[index,];
dataTrainValidation=dataTrain1[-index,];
```
## setting training control parameters
```{r,echo=TRUE, results='hide', message=FALSE}
tc=trainControl(method='repeatedcv', number = 2)
```


## Training with SVM
training with support vectors machine
```{r,echo=TRUE, results='hide', message=FALSE}
modfSVM<-train(classe~., data = dataTrain, method='svmLinear', trControl=tc, preProcess=c('center','scale'))
```
SVM model accuracy
```{r,echo=TRUE, message=FALSE}
modfSVM
confusionMatrix(dataTrainValidation$classe,predict(modfSVM,dataTrainValidation[,-53]))

```
## Training with k nearest neighbors
training with k nearest neighbors
```{r,echo=TRUE, results='hide', message=FALSE}
modfKNN<-train(classe~., data = dataTrain, method='knn', trControl=tc, preProcess=c('center','scale'))

```
k nearest neighbors model accuracy
```{r,echo=TRUE, message=FALSE}
modfKNN
confusionMatrix(dataTrainValidation$classe,predict(modfKNN,dataTrainValidation[,-53]))

```
## Training with random forest
training with random forest
```{r rf,echo=TRUE, results='hide', message=FALSE}
modfRF<-train(classe~., data = dataTrain, method='rf', trControl=tc, preProcess=c('center','scale'))


```
random forest model accuracy
```{r rfAc,echo=TRUE, message=FALSE}
modfRF
confusionMatrix(dataTrainValidation$classe,predict(modfRF,dataTrainValidation[,-53]))
```

## Training with partition trees
training with partition trees
```{r rpart,echo=TRUE, results='hide', message=FALSE}
modfRPART<-train(classe~., data = dataTrain, method='rpart', trControl=tc, preProcess=c('center','scale'),tuneGrid=data.frame(cp=0.0001))
```
partition trees model accuracy
```{r rpartAc,echo=TRUE, message=FALSE}
modfRPART
confusionMatrix(dataTrainValidation$classe,predict(modfRPART,dataTrainValidation[,-53]))
```
## Training with gradient boosting machine
training with gradient boosting machine
```{r gbm,echo=TRUE, results='hide', message=FALSE}
modfGBM<-train(classe~., data = dataTrain, method='gbm', trControl=tc, preProcess=c('center','scale'))

```
gradient boosting machine model accuracy
```{r gmbAc,echo=TRUE, message=FALSE}
modfGBM
confusionMatrix(dataTrainValidation$classe,predict(modfGBM,dataTrainValidation[,-53]))
```
## Conclusion
Comparing the accuracies of all five models we can see that "random forest" has the best results. Therefore we can apply it on the test data.
```{r testdata, echo=TRUE}
predict(modfRF,TestData)
```
